Clustering Beat-Chroma Patterns\\ in a Large Music Database

### 2010-03-24 Paper/Paper.tex
made math prettier

@@

The representation resulting from this process is invariant to both
the key and meter of the original song.  This enables the study of
broad harmonic patterns found throughout the data, without regard for
the specific musical context.
%
% In the context of clustering, this enables us to study harmonic
% pattern usage without maintaining clusters for the specific context a
% particular patterns occurs in,
In the context of clustering this avoids \eg obtaining separate
clusters for every major triad in duple or triple meter.
% Also, this implies that the pitch envelope is important rather than
% the actual pitches.
Something about why we're not interested in specific chords?

@@

We use an online version of the vector quantization algorithm
\cite{Gersho1991}, to cluster the beat-chroma patches described in the
previous section.
% This can also be seen as an online K-means.
For each sample from the data, the algorithm finds the closest cluster
in the codebook and updates the cluster centroid (codeword) to be
closer to the sample based on some learning rate $l$.
The clusters are updated as each data point is seen, as opposed to
once per iteration in the standard k-means algorithm.
The details are explained in Algorithm \ref{algo:vq}.
%
As in standard k-means clustering, the codebook is initialized by
choosing $K$ random points from our dataset.
%
Note that this algorithm, although not optimal, scales linearly with
the number of patches seen and can be interrupted at any time
while still resulting in an updated codebook. %not that this is important...
% The setting of the learning rate might become trickier, but this is
% not fundamentally a problem.

\begin{algorithm}
%\caption{Pseudocode of Vector Quantization}
\begin{algorithmic}
\STATE$l$ learning rate
\STATE$\{P_n\}$ set of patches
\STATE$\{C_k\}$ codebook of $K$ codes
%\STATE $m \leftarrow b$
\REQUIRE $0 < l \leq 1$
\FOR{$nIters$}
\FOR{$p \in \{P_n\}$}
\STATE$c \leftarrow min_{c \in C_k} dist(p,c)$
\STATE$c \leftarrow c + (p - c) * l$
\ENDFOR
\ENDFOR
\RETURN $\{C_k\}$
\caption{\small{Pseudocode of Online Vector Quantization. Note that we can 
replace the number of iterations by a threshold on the distortion over some 
test set.}
\label{algo:vq}}
\end{algorithmic}
\end{algorithm}

@@

As a distance measure (or distortion measure) between patterns, we use
the average squared euclidean distance. In details, if a pattern $p1$
is composed of elements $p1_{i,j}$, and similarly for a pattern $p2$,
the distance between $p1$ and $p2$ is:
\begin{eqnarray}
  dist(p1,p2) = \sum_{i,j} (p1_{i,j} - p2_{i,j})^2 / size(p1)  \label{eq:dist}
\end{eqnarray}
We assume $p1$ and $p2$ have the same size.

>>>>

\DeclareMathOperator*{\dist}{dist}
\DeclareMathOperator*{\size}{size}

@@

The representation resulting from this process is invariant to both
the key and meter of the original song.  This enables the study of
broad harmonic patterns found throughout the data, without regard for
the specific musical context.
%
% In the context of clustering, this enables us to study harmonic
% pattern usage without maintaining clusters for the specific context a
% particular patterns occurs in,
In the context of clustering this avoids \eg obtaining separate
clusters for every major triad in duple or triple meter.
% Also, this implies that the pitch envelope is important rather than
% the actual pitches.
FIXME: Something about why we're not interested in specific chords?

@@

We use an online version of the vector quantization algorithm
\cite{Gersho1991}, to cluster the beat-chroma patches described in the
previous section.
% This can also be seen as an online K-means.
For each sample from the data, the algorithm finds the closest cluster
in the codebook and updates the cluster centroid (codeword) to be
closer to the sample based on some learning rate $\ell$.
The clusters are updated as each data point is seen, as opposed to
once per iteration in the standard k-means algorithm.
The details are explained in Algorithm \ref{algo:vq}.
%
As in standard k-means clustering, the codebook is initialized by
choosing $K$ random points from our dataset.
%
Note that this algorithm, although not optimal, scales linearly with
the number of patches seen and can be interrupted at any time
while still resulting in an updated codebook. %not that this is important...
% The setting of the learning rate might become trickier, but this is
% not fundamentally a problem.

\begin{algorithm}
%\caption{Pseudocode of Vector Quantization}
\begin{algorithmic}
\STATE$\ell$ learning rate
\STATE$\{P_n\}$ set of patches
\STATE$\{C_k\}$ codebook of $K$ codes
%\STATE $m \leftarrow b$
\REQUIRE $0 < \ell \leq 1$
\FOR{$nIters$}
\FOR{$p \in \{P_n\}$}
\STATE$c \leftarrow min_{c \in C_k} \dist(p,c)$
\STATE$c \leftarrow c + (p - c) * \ell$
\ENDFOR
\ENDFOR
\RETURN $\{C_k\}$
\caption{\small{Pseudocode for the online vector quantization
    algorithm. Note that we can replace the number of iterations by a
    threshold on the distortion over some test set.}
\label{algo:vq}}
\end{algorithmic}
\end{algorithm}

@@

As a distance measure (or distortion measure) between patterns, we use
the average squared euclidean distance. In details, if a pattern $p_1$
is composed of elements $p_{1ij}$, and similarly for a pattern $p_2$,
the distance between $p_1$ and $p_2$ is:
\begin{eqnarray}
  \dist(p_1,p_2) = \sum_{i,j} \frac{(p_{1ij} - p_{2ij})^2}{\size(p_1)}
  \label{eq:dist}
\end{eqnarray}
We assume $p_1$ and $p_2$ have the same size.
